### 2015-01-07 Papers/Prototype_1_Method/article_table.tex

\begin{table}
\begin{center}
\begin{tabular}{||c|| c c c c ||}
\hline\hline
{\bf Mesh} & {\bf Number of}        & {\bf Number of} & {\bf Element}  & {\bf Time-Step}    \\
           & {\bf Elements}$^{\star}$&{\bf Nodes}      & {\bf Size}     & {\bf Size}  \\
\hline
1          & 68 / 3 / 3                & 47              & 0.30           & 5.0$\times$10$^{-2}$ \\
2          & 538 / 12  / 28             & 304             & 0.10           & 1.667$\times$10$^{-2}$ \\
3          & 2316 / 46  / 115            & 1229            & 0.05           & 8.333$\times$10$^{-3}$ \\
4          & 6410 / 130  / 320           & 3324            & 0.03           & 5.0$\times$10$^{-3}$ \\
5          & 57650 / 1170 / 2886          & 29176           & 0.01           & 1.667$\times$10$^{-3}$ \\
\hline\hline
\end{tabular}
\caption{Advection problem test-cases: Meshes used in the simulations. $\left(^{\star}\right)$ refers to the number of elements in the domain and within the Gaussian plume and step-function. \label{table:meshplume}}
\end{center}
\end{table}