### 2018-03-13 Papers/AppliedMaths2018/article_figure1.tex

%%%%
%%%%  FIGURE 
%%%%
\begin{figure}[h!]
   \vbox{
     \hbox{\includegraphics[width=\columnwidth,clip]{./Pics/PermField_Saturation.png}}
     \hbox{\hspace{4cm}(a)\hspace{7.8cm}(b)}}
        \caption{(a) Permeability distribution and (b) phase saturation showing preferential flow pathways.}
        \label{fig:PermField_Saturation}  
\end{figure}
\clearpage