### 2018-03-13 Papers/Prototype_Merged/article.tex

%%
%% Copyright 2007, 2008, 2009 Elsevier Ltd
%%
%% This file is part of the 'Elsarticle Bundle'.
%% ---------------------------------------------
%%
%% It may be distributed under the conditions of the LaTeX Project Public
%% License, either version 1.2 of this license or (at your option) any
%% later version.  The latest version of this license is in
%%    http://www.latex-project.org/lppl.txt
%% and version 1.2 or later is part of all distributions of LaTeX
%% version 1999/12/01 or later c.
%%
%% The list of all files belonging to the 'Elsarticle Bundle' is
%% given in the file `manifest.txt'.
%%

%% Template article for Elsevier's document class `elsarticle'
%% with harvard style bibliographic references
%% SP 2008/03/01
%%
%%
%%
%% $Id: elsarticle-template-harv.tex 4 2009-10-24 08:22:58Z rishi $
%%
%%
\documentclass[preprint,authoryear,12pt]{elsarticle}

%% Use the option review to obtain double line spacing
%% \documentclass[authoryear,preprint,review,12pt]{elsarticle}

%% Use the options 1p,twocolumn; 3p; 3p,twocolumn; 5p; or 5p,twocolumn
%% for a journal layout:
%% \documentclass[final,authoryear,1p,times]{elsarticle}
%% \documentclass[final,authoryear,1p,times,twocolumn]{elsarticle}
%% \documentclass[final,authoryear,3p,times]{elsarticle}
%% \documentclass[final,authoryear,3p,times,twocolumn]{elsarticle}
%% \documentclass[final,authoryear,5p,times]{elsarticle}
%% \documentclass[final,authoryear,5p,times,twocolumn]{elsarticle}

%% if you use PostScript figures in your article
%% use the graphics package for simple commands
%% \usepackage{graphics}
%\usepackage{cases}
%% or use the graphicx package for more complicated commands
\usepackage{graphicx}
%% or use the epsfig package if you prefer to use the old commands
\usepackage{epsfig}

\usepackage{comment}

\usepackage{epstopdf}

\usepackage{subfig}

\usepackage{bm}
\usepackage{hyperref,url}
\hypersetup{colorlinks=true, urlcolor=blue, linkcolor=blue, citecolor=red}

%\usepackage[obeyFinal]{todonotes}
\usepackage{todonotes}
\newcommand{\jrp}[1]{\todo[color=blue!30, size=\small]{JRP: #1}}
\newcommand{\JGnote}[1]{\fbox{\parbox{\textwidth}{ \color{blue} JG: #1}}}

%% The amssymb package provides various useful mathematical symbols
\usepackage{amssymb,amsmath,array}
%% The amsthm package provides extended theorem environments
%% \usepackage{amsthm}

%% The lineno packages adds line numbers. Start line numbering with
%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
%% for the whole article with \linenumbers after \end{frontmatter}.
%% \usepackage{lineno}

%% natbib.sty is loaded by default. However, natbib options can be
%% provided with \biboptions{...} command. Following options are
%% valid:

%%   round  -  round parentheses are used (default)
%%   square -  square brackets are used   [option]
%%   curly  -  curly braces are used      {option}
%%   angle  -  angle brackets are used    <option>
%%   semicolon  -  multiple citations separated by semi-colon (default)
%%   colon  - same as semicolon, an earlier confusion
%%   comma  -  separated by comma
%%   authoryear - selects author-year citations (default)
%%   numbers-  selects numerical citations
%%   super  -  numerical citations as superscripts
%%   sort   -  sorts multiple citations according to order in ref. list
%%   sort&compress   -  like sort, but also compresses numerical citations
%%   compress - compresses without sorting
%%   longnamesfirst  -  makes first citation full author list
%%
%% \biboptions{longnamesfirst,comma}

% \biboptions{}

%\journal{Computer Methods in Applied Mechanics and Engineering}
\journal{Comput Meth Appl Mech Eng}
\newcommand{\frc}{\displaystyle\frac}
\newcommand{\PN}[2][error]{P$_{#1}$DG-P$_{#2}$}
\newcommand{\PNDG}[2][error]{P$_{#1}$DG-P$_{#2}$DG}

\begin{document}

\begin{frontmatter}

%% Title, authors and addresses

%% use the tnoteref command within \title for footnotes;
%% use the tnotetext command for the associated footnote;
%% use the fnref command within \author or \address for footnotes;
%% use the fntext command for the associated footnote;
%% use the corref command within \author for corresponding author footnotes;
%% use the cortext command for the associated footnote;
%% use the ead command for the email address,
%% and the form \ead[url] for the home page:
%%
%% \title{Title\tnoteref{label1}}
%% \tnotetext[label1]{}
%% \author{Name\corref{cor1}\fnref{label2}}
%% \ead{email address}
%% \ead[url]{home page}
%% \fntext[label2]{}
%% \cortext[cor1]{}
%% \address{Address\fnref{label3}}
%% \fntext[label3]{}

\title{A Force-Balanced Control Volume Finite Element Method for Multiphase Porous Media Flow Modelling}

%% use optional labels to link authors explicitly to addresses:
\author[UoA,NORMS]{J. L. M. A. Gomes}\corref{cor1}\ead{jefferson.gomes@abdn.ac.uk} \author[IC,NORMS]{D. Pavlidis} 
\author[IC,NORMS]{P. Salinas} \author[IC,NORMS]{Z. Xie}
\author[IC]{J. R. Percival} \author[NSW]{P. Mostaghimi} 
\author[IC,NORMS]{C. C. Pain} \author[NORMS]{M. D. Jackson}
%\author[IC]{M. Blunt}

\cortext[cor1]{Corresponding author.}
\address[UoA]{Environmental and Industrial Fluid Mechanics Group, School of Engineering, University of Aberdeen, UK}
\address[IC]{Applied Modelling and Computation Group, Department of Earth Science and Engineering, Imperial College London, UK}
\address[NSW]{School of Petroleum Engineering, University of New South Wales, Australia}
\address[NORMS]{Novel Reservoir Modelling and Simulation Group, Department of Earth Science and Engineering, Imperial College London, UK}

\begin{abstract}
  A novel control volume finite element method for simulating
  subsurface multiphase flows is presented here. The approach has two
  major strengths. First, it uses a dual consistent pressure-velocity
  representation in control volume and finite element spaces which
  ensures local mass conservation. Second, it uses a family of
  finite element types that ensure the multiphase Darcy equations can be
  exactly enforced. The underlying mass conservation equations are
  solved in control volume space whereas finite elements are used to
  obtain high-order fluxes on control volume boundaries. The novelty
  of the method lies in (a) permitting both continuous and
  discontinuous description of pressures and saturations between
  elements; (b) the use of arbitrarily high-order polynomial
  representation for pressure and velocity within the mixed
  formulation and (c) the use of high-order flux-limited methods to
  avoid introducing non-physical oscillations while achieving
  high-order time accuracy where and when possible. The model is
  initially evaluated and then applied to a series of test cases using
  structured and unstructured (triangular/ tetrahedral) meshes, as
  well as heterogeneous problems with large permeability
  ratios. Numerical results are in good agreement with analytically
  obtained solutions.
\end{abstract}

\begin{keyword} %% keywords here, in the form: keyword \sep keyword
CVFEM mixed formulation \sep Discontinuous Galerkin  \sep Multiphase flows \sep Porous media.

\end{keyword}

\end{frontmatter}

\section{Introduction}
Numerical modelling of multiphase flow in porous media has
applications in a wide range of disciplines, including hydrocarbon and
groundwater production, safety assessment of deep geological disposal
of radioactive waste, and CO$_{\text{2}}$ migration and trapping in
carbon capture and storage ~\citep{chen_2006, aiea_1999, pruess_1990c,
  jiang_2011}. Such geological and geophysical applications often
include complex rock geometries with irregular (and often internal)
boundaries between rock types with contrasting porosity and
permeability.  Accurately capturing these complex geometries is a key
challenge when simulating such flows.

Finite difference methods (FDM) have been extensively used in
modelling and simulation of fluid flows in porous media
\citep{aziz_1986, chen_1997, chen_2005}.  However, they are strongly
dependent on mesh quality and orientation, and cannot easily represent
complex geometries. In addition, FDM can produce excessive numerical
dispersion in heterogeneous porous media \citep{chavent_1986}. The
geometrical flexibility of finite element methods (FEM) has been shown
to overcome these deficiencies. Among FEM-based formulations for
porous media, the control volume finite element method (CVFEM) has
become increasingly popular as it can guarantee local mass
conservation and high-order numerical accuracy, as well as being able
to use tetrahedral, geometry-conforming elements \citep{forsyth_1990,
  cordazzo_2004, geiger_2004, hurtado_2007}. \citet{huber_2000}
demonstrated that the vertex-centred, finite volume-finite element
method (or Box scheme) can achieve similar goals \citep[see
  also][]{helmig_1997}.

\citet{durlofsky_1993,durlofsky_1994} compared the performance of a
mixed FEM formulation (i.e. piece-wise constant in pressure and
piece-wise linear in velocity) and the classical CVFEM for single
phase Darcy flows. He concluded that the latter is more
computationally efficient and numerically accurate than the former for
problems involving a (relatively) small number of degrees of
freedom. Moreover, for complex heterogeneous problems with sharp
changes in material properties, mixed FEM formulations often lead to
physically unrealistic solutions.  \citet{cumming_2011} demonstrated
that a CVFEM-based discretisation could be used to solve the Richards
equation (coupled mass conservation and Darcy equations) in
heterogeneous porous media with relatively small computational
overhead, compared with traditional, coupled velocity-pressure based
formulations.  Mass balance was enforced as described by
\citet{kirkland_1992}~\citep[see
  also][]{forsyth_1990,cumming_phd2012}.  The efficiency and accuracy
of CVFEM for porous media flows has motivated our development of a
CVFEM-based discretisation here.  However, CVFEM often requires high
resolution meshes in regions where material properties vary abruptly,
such as permeability contrasts at (for example) fracture-matrix
interfaces, or boundaries between different rock types. Control volume
(CV) boundaries span finite elements where material properties are
defined; therefore, some average value of the permeability must be
calculated across CV interfaces. This often leads to excessive
numerical dispersion, especially in highly heterogeneous media
\citep{nick_2011b, nick_2011a}.

Discontinuity-capturing schemes~\citep[e.g. shock waves, contact
  surface or material discontinuities --
  see][]{brooks_1982,tezduyar_1986} were originally developed to
resolve sharp changes in solution fields such as velocity or
saturation. Related discontinuity-capturing schemes include the
discontinuous-Galerkin FEM (DGFEM) scheme in which continuity of the
solution is not explicitly enforced, allowing sharp changes in the
solution fields to be captured. The DG formulation is a stabilised
finite element method that is locally conservative and designed to
achieve high-order accuracy. DGFEM solution fields are allowed to be
discontinuous at the element faces; thus, the solution is able to
handle discontinuities in material properties at internal
boundaries. In addition, DGFEM is well-suited to deal with interface
problems by incorporating specially-designed interface fluxes. These
properties have attracted the attention of the porous media flow
community over the past 15 years
\citep[see][]{riviere_2000,riviere_2002,bastian_2002} and are utilised
here.

In this paper, a novel CVFEM formulation which is conservative and
consistent is presented. The continuity equations are embedded into
the pressure equation to enforce mass conservation and the exact force
balance. The formulation employs an implicit algorithm with respect to
time that is less restrictive than the
implicit-pressure-explicit-saturation (IMPES) scheme often adopted in
porous media flow problems \citep[e.g. see][]{aziz_1986,
  geiger_2004}. There are two key novel aspects in our approach. The
first is that it uses a dual consistent CVFEM representation that is
embedded in the \PN[n]{m}[DG] family of triangular and tetrahedral
finite element pairs \citep[see][]{kaisu}. For this element type, the
velocity field is represented by $n^{\rm th}$-order polynomials that
are discontinuous across elements, while the pressure field is
represented by $m^{\rm th}$-order polynomials that may be continuous
(termed P$_n$DG-P$_m$) or discontinuous (termed P$_n$DG-P$_m$DG)
across elements. In particular, here the P$_n$DG-P$_{n+1}$ element
pair, which allows the velocity to exactly represent the pressure
gradients in the flow solution for homogeneous material properties, is
demonstrated. The \PNDG[n+1]{n} element pair, which has similar
properties to the \PN[n]{n+1} element pair, but allows a
representation in which pressure, saturation and other solution
variables are discontinuous across finite element boundaries, is also
demonstrated.  This solves the long-standing problem described above
with respect to the use of traditional CVFEM methods to capture sharp
changes in material properties. The second key novelty is the velocity
representation that is described in detail in
Section~\ref{overlapping_method_section} and can achieve an exact
representation of the multiphase Darcy force balance equations.

The remainder of this paper is organised as follows. The governing
equations and numerical methods employed to solve them are presented
in Section~\ref{overlapping_method_section}. Model set-up and results
including comparison against reference data for a series of benchmark
problems are given in Sections~\ref{setup} and \ref{res},
respectively. Finally, some concluding remarks are presented in
Section~\ref{conc}.

\section{Governing equations and Numerical Method}
\label{overlapping_method_section}

\subsection{A Novel Representation of Multiphase Darcy Flows}
Darcy's law for immiscible multiphase flow may be written in the form:
\begin{equation}\label{e:darcy_eqn}
  \mathbf{q}_{\alpha} =
  -\frac{\mathcal{K}_{{r}_\alpha}\mathbf{K}}{\mu_{\alpha}}\left(
  \nabla p_{\alpha} - {\mathbf{s}_{u}}_{\alpha} \right),
\end{equation}
where $\mathbf{q}_{\alpha}$ is the $\alpha$-th phase Darcy velocity,
$\mathbf{K}$ is the absolute permeability tensor of the porous medium,
$\mathcal{K}_{{r}_\alpha}\left(S_{\alpha}\right)$ is the phase
relative permeability, which is a function of the phase saturation
$S_{\alpha}\left(\mathbf{r},t\right)$. $\mu_{\alpha}$, $p_{\alpha}$
and $\mathbf{s}_{{u}_\alpha}$ are the phase dynamic viscosity,
pressure and source term, which may include gravity and capillary
pressure, respectively.

Introducing a saturation-weighted Darcy velocity defined as
$\mathbf{u}_\alpha= \mathbf{q}_\alpha/S_\alpha$, then
Eqn.~\ref{e:darcy_eqn} may be rewritten as:
\begin{equation}
  \mathbf{v}_\alpha={\underline {\underline \sigma}}_{\alpha}
  \mathbf{u}_{\alpha} = - \nabla p_{\alpha} +
         {\mathbf{s}_{u}}_{\alpha},
  \label{force-bal}
\end{equation}
where ${\underline {\underline \sigma}}_{\alpha}=\mu_\alpha S_\alpha
\left(\mathcal{K}_{{r}_\alpha}\mathbf{K}\right)^{-1}$ represents the
implicit linearisation of the viscous frictional forces. The force per
unit volume $\mathbf{v}_\alpha$, defined as ${\underline {\underline
    \sigma}}_{\alpha} \mathbf{u}_\alpha$, is used as a prognostic
variable in this approach, as explained in
Section~\ref{Section:Force_density}.

In order to discretise Eqn.~\ref{force-bal}, a finite element
representation for $\mathbf{v}_\alpha$ and $p$ is assumed, expressed
in terms of their FE basis functions $Q_{j}$ and $P_{j}$,
respectively, as:
\begin{equation}
  \mathbf{v}_\alpha(\bm{r},t) = \sum\limits_{j=1}^{\mathcal{N}_u}
  Q_{j}(\bm{r})\mathbf{v}_{\alpha,j}(t) \;\;\;\;\text{ and } \;\;\;\;
  p(\bm{r},t) = \sum\limits_{j=1}^{\mathcal{N}_p}
  P_{j}(\bm{r})p_{j}(t).
\end{equation} 
Here $\mathcal{N}_{u}$ and $\mathcal{N}_{p}$ are the total number of
degrees of freedom for the FE force and pressure representations. Each
component of the weak form of the force balance (Eqn.~\ref{force-bal})
is tested with the $\mathbf{v}_\alpha$ basis function space to obtain:
\begin{eqnarray}
  \sum\limits_{E} \left. \int\limits_{\Omega_E} { {Q}}_i \left (
             {\mathbf v}_\alpha + \nabla p -{\mathbf s}_{u_\alpha}
             \right) dV \right. + \displaystyle \oint_{\Gamma_{E}}
                   {Q}_i {\mathbf n} \left(p - \tilde{p}\right)
                   d\Gamma + \nonumber \\ \oint_{\Gamma_{\Omega}} {
                     Q}_i {\mathbf n} \left(p - p_\text{bc}\right) d\Gamma
                   = \bm{0},
                   \label{force-semi-disc} 
\end{eqnarray} 
where $\Omega_E$ and $\Gamma_{E}$ are the volume and boundary of
element $E$, respectively, and $\Gamma_{\Omega}$ is the boundary of
the computational domain. The numerical pressure $\tilde{p}$ appearing
in the jump condition (second term in Eqn.~\ref{force-semi-disc}) is
the arithmetic mean of the potentially discontinuous pressure across
the element $E$. This term vanishes when a continuous formulation is
used to discretise the pressure field. The last term in
Eqn.~\ref{force-semi-disc} is used to weakly enforce the pressure
level to $p_\text{bc}$ on a computational domain boundary.

In matrix form, Eqn.~\ref{force-semi-disc} is:
\begin{equation}
  {\mathbf M} \underline {\mathbf v} = -{\mathbf C} \underline
  {\mathbf p} + \underline {\mathbf
    s}_{u}, \label{force-balance-matrix-form}
\end{equation}
where $\underline{\bf v}$ and $\underline{\bf p}$ solution-vectors are
defined as:
\begin{eqnarray}
  &&\underline{{\bf v}} = \left( 
  \left( {v_x},{v_y},{v_z} \right)  _{1,1},
  \left( {v_x},{v_y},{v_z} \right)  _{2,1},\ldots,
  \left( {v_x},{v_y},{v_z} \right)  _{\mathcal{N}_\alpha,\mathcal{N}_u}
  \right)^{T}
  \;\text{and} \nonumber \\ &&
  \underline{{\bf p}} = \left(p_1, p_2, p_3, ...,
  p_{\mathcal{N}_p}\right)^T. \nonumber \nonumber
\end{eqnarray}
$\mathcal{N}_{\alpha}$ is the number of phases and $v_x$, $v_y$ and
$v_z$ are the components associated with the $x$, $y$ and $z$
dimensions, respectively. Finally, the mass matrix ${\mathbf M}$,
gradient matrix ${\mathbf C}$ and source vector $\underline {\mathbf
  s}_{u}$ are defined as:
\begin{eqnarray}
  && \left[{\mathbf M}^{(i,j)}\right]_{k,l} = \int_{\Omega} Q_{k}
  \left[\underline{\underline{\sigma}}_{\alpha}\right]_{i,j} {Q}_{l}
  dV, \nonumber \\ && {\left[\bm{C}^{(i,i)}\right]_{k,l} =
    \int_{\Omega} Q_{k} \frac{\partial P_{l}}{\partial r_i} dV -
    \oint_{\Gamma_{E}} \displaystyle\frac{1}{2} Q_{k} n_i^{(E,l)}
    P_{l} d\Gamma - \oint_{\Gamma_{\Omega}} \displaystyle Q_{k} n_i
    P_{l} d\Gamma}, \nonumber \\ &&
  {\left[\bm{s}_{u}^{(i,i)}\right]_{k} = \int_{\Omega} Q_k}
  \left[{\mathbf s}_{u_\alpha} \right]_i dV -
  \oint_{\Gamma_{\Omega}}\displaystyle Q_{k} n_i p_\text{bc} d\Gamma,
  \nonumber
\end{eqnarray}
where $i$ and $j$ are dimensions and $k$ and $l$ are the degrees of
freedom of element $E$. Here, $\bm{n} = \left(n_{1}, n_{2},
n_{3}\right)^{T}$ is the outward-pointing normal vector of the domain
$\Omega$. $\bm{n}^{(E,l)}$ is the normal to the boundary of element
$E$ and is outward-pointing if $P_{l}$ takes support on $E$, and
inward-pointing otherwise.

The saturations are computed in CV space, whereas the permeability
tensor $\left(\mathbf{K}\right)$ is assumed piece-wise constant in FE
space. Thus the viscous-friction damping tensor $ {\underline
  {\underline \sigma}}_{\alpha}$ is piece-wise constant within the
sub-control volumes defined through intersections between the elements
and control volumes (see Fig.~\ref{fem_cv_represent_a}).

\subsection{Control-Volume Finite element types}
In the formulation presented here, pressure, velocity, permeability
and porosity are represented FE-wise, but saturation, relative
permeability and fluid properties such as viscosity and density are
represented CV-wise. Fig.~\ref{fem_cv_represent_a} displays the two
families of element types presented in this
paper. Fig.~\ref{fem_cv_represent_a} (a) shows the P$_{n}$DG-P$_{n+1}$
element type, with $n = 1$, in two dimensions (2D). Here the velocity
field is linear, discontinuous (between elements), the pressure field
has a quadratic, continuous (between elements), representation and the
CV's span various elements. Fig.~\ref{fem_cv_represent_a} (b) shows
the P$_{n+1}$DG-P$_{n}$DG element type, with $n = 1$, also in 2D. Here
the velocity field is quadratic, discontinuous (between elements), the
pressure field has a linear, discontinuous (between elements),
representation and the CV's do not span elements. Note that in the
latter element pair, CV's have piece-wise constant permeability and
porosity values.

\begin{figure}[h!]
\subfloat[P$_1$DG-P$_2$]{\includegraphics[width=.5\textwidth]{diagrams/p1dgp2-cont-sat}}
\subfloat[P$_2$DG-P$_1$DG]{\includegraphics[width=.5\textwidth]{diagrams/p2dgp1-dg-sat}}
\caption{2D representation of the element pairs presented in this
  work. Shaded areas denote control volumes (in which saturation is
  stored), black points represent the pressure nodes and white points
  the velocity nodes.}
    \label{fem_cv_represent_a}
\end{figure}

\subsection{Finite element representation of velocity} \label{Section:Force_density}
From Eqn.~\ref{e:darcy_eqn}, it can be seen that the velocity field
depends, among other fields, on the absolute permeability, which is
FE-wise, and the relative permeability, which is CV-wise. This means
that the velocity representation is different in each CV. However, the
quantity $\mathbf{v}_\alpha$ is homogeneous within an element and this
is what is stored using FE. To obtain the velocity
$\mathbf{u}_{\alpha}$ in a particular CV within a FE,
Eqn.~\ref{force-bal} is used.

\subsection{Saturation and Global Mass Conservative Equations} \label{Section:Saturation_Global}
Here, the saturation conservation equations are discretised and the
global mass balance equation is derived. In this work, fluids are
assumed incompressible. The saturation equation can be written as:
\begin{equation}
  \phi\displaystyle\frac{\partial S_{\alpha} }{\partial t} + \nabla
  \cdot \left( {\mathbf u}_{\alpha} S_{\alpha}\right) =
  s_{cty,\alpha},
  \label{saturation_equation}
\end{equation}
where $\phi$ is the porosity and $s_{cty}$ is a source
term. Eqn.~\ref{saturation_equation} is discretised in space by
testing with CV basis functions $M_{i}$ and with the $\theta$-method
in time:
\begin{eqnarray}
  \int_{\Omega} M_{i} \displaystyle\frac{\phi \left({S_{\alpha
        i}^{n+1}}-{S_{\alpha i}^{n}}\right)}{\Delta t} dV + \nonumber
  \\ \oint_{\Gamma_{CV_{i}}} \left[\theta^{n+\frac{1}{2}} {\mathbf n}
    \cdot {\mathbf u}_{\alpha}^{n+1} S_{\alpha}^{n+1} +
    \left(1-\theta^{n+\frac{1}{2}}\right) {\mathbf n} \cdot {\mathbf
      u}_{\alpha}^{n} S_{\alpha}^{n} \right]d\Gamma = \int_{\Omega}
  M_{i} {s_{cty,\alpha}^{n+\theta}} dV,
  \label{detail-sat-eqn-k}
\end{eqnarray}
where $\mathbf{n}$ is the outward-pointing unit normal vector to the
surface ($\Gamma_{CV_{i}}$) of the control volume $i$ and $n$ is the
current time level. More details on the numerical methods used to
solve Eqn.~\ref{detail-sat-eqn-k} can be found in
\citet{pavlidis_2013b}.

The global continuity equation is obtained by summing
Eqn.~\ref{detail-sat-eqn-k} over all phases:
\begin{eqnarray}
  \sum\limits_{\alpha=1}^{\mathcal{N}_{\alpha}} \left\lbrace
  \int_{\Omega} M_{i} \displaystyle\frac{\phi\left({S_{\alpha
        i}^{n+1}}-{S_{\alpha i}^{n}}\right) } {\Delta t} dV +
  \right. \nonumber \\ \left. \displaystyle\oint_{{\Gamma_{CV}}_{i}}
  \left[\theta^{n+\frac{1}{2}} {\mathbf n} \cdot {\mathbf
      u}_{\alpha}^{n+1} S_{\alpha}^{n+1} +
    \left(1-\theta^{n+\frac{1}{2}}\right) {\mathbf n} \cdot {\mathbf
      u}_{\alpha}^{n} S_{\alpha}^{n} \right] \,d\Gamma -
  \right. \nonumber \\ \left. \displaystyle\int_{\Omega} M_{i}
  s_{cty,\alpha}^{n+\theta}\, dV \right\rbrace = 0.
           \label{detail-sat-eqn-k-sum}
\end{eqnarray}
Eq.~\ref{detail-sat-eqn-k-sum} is also bounded by the constraint:
\begin{equation}
  \sum\limits_{\alpha=1}^{\mathcal{N}_{\alpha}} {S_{\alpha}^{n}}_{i} =
  1, \quad \forall n,
\end{equation}
Solving for ${\mathbf u}^{n+1}$ in matrix form,
Eqn.~\ref{detail-sat-eqn-k-sum} becomes:
\begin{equation}
  {\mathbf B}^T \underline{\bf v}^{n+1} = \underline{\bf s}_{p}.
  \label{glob-cty-matrix}
\end{equation}

\subsection{Velocity at Control Volume Interfaces} \label{opt-up} 
The velocity to be used at the interface between CV's in the
saturation conservation (Eqn.~\ref{detail-sat-eqn-k}) and global
continuity (Eqn.~\ref{detail-sat-eqn-k-sum}) equations needs to be
determined. On the CV faces, there is no information about the flow
direction as the velocity is discontinuous at the CV boundaries. The
discontinuities occur between (a) control volumes within each FE and
(b) between elements when using the P$_{n+1}$DG-P$_{n}$DG pair.

To calculate the velocity across CV's within a FE, an average velocity
at the interface of control volumes {\it i} and {\it j} is defined as:
\begin{equation}
  {\tilde{\bf u}}_\alpha = \frac{1}{2} \left[ {{\bf u}_\alpha}_i +
    {{\bf u}_\alpha}_j \right].
\end{equation} 
From this, the interface velocities at either side of the interface
can be obtained from:
\begin{equation}
  {\tilde{\bf u}}_{\alpha k} = \underline{\underline{\sigma}}_{\alpha
    k}^{-1}{\tilde{{\bf u}}}_{\alpha}, \;\;\;\; k=\{i,j\}.
  \label{two-vels}
\end{equation} 
These velocities have the same direction and 
differ only in magnitude, so using them to define an upwind
direction is not ambiguous.

If an upwind method for calculating the interface velocity is applied
then the interface velocity becomes:
\begin{equation}
  {\tilde{\bf u}}_{\alpha} = \tilde{\bf u}_{\alpha k},
\end{equation}
where $k=i$, if ${\bf n}\cdot{\bf u}_{\alpha}>0$ (CV {\it i} outgoing
information) and $k=j$, if ${\bf n}\cdot{\bf u}_{\alpha}<0$ (CV {\it
  i} incoming information). However, upwind methods yield dissipative
solutions. In order to obtain a high-order approximation of the
velocity at the interface, the saturation at the interface
$\tilde{S}_{\alpha}$ is calculated using a finite element
representation of the saturation following the upwind direction.
$\tilde \sigma_{\alpha}$ can then be calculated using second-order
Taylor series:
\begin{small}
\begin{equation}
  \underline{\underline{\tilde{\sigma}}}_{\alpha} =
  \underline{\underline{\sigma}}_{\alpha k} +
  \left(\tilde{S}_{\alpha}-S_{\alpha k}\right) \left(\frc{\partial
    \underline{\underline{\sigma}}_{\alpha}}{\partial
    S_{\alpha}}\right)_{k} +
  \frc{1}{2}\left(\tilde{S}_{\alpha}-S_{\alpha k}\right)^{2} \left[
    \frc{ \left(\frc{\partial
        \underline{\underline{\sigma}}_{\alpha}}{\partial
        S_{\alpha}}\right)_{k} - \left(\frc{\partial
        \underline{\underline{\sigma}}_{\alpha}}{\partial
        S_{\alpha}}\right)_{l} } { \left(S_{\alpha k}-S_{\alpha
        l}\right) } \right],
  \label{sigma-out}
\end{equation}
\end{small}
where $k=i$ and $l=j$, if ${\bf n}\cdot{\bf u}_{\alpha}>0$ and $k=j$
and $l=i$, if ${\bf n}\cdot{\bf u}_{\alpha}<0$. The final interface
velocity is obtained from:
\begin{equation}
  {\tilde{\bf u}_{\alpha}} =
  \underline{\underline{\tilde{\sigma}}}_{\alpha}^{-1}
  \left[\displaystyle\frac{1}{2} \left( {{\bf u}_\alpha}_i + {{\bf
        u}_\alpha}_j \right)\right],
  \label{mean_int_vel}  
\end{equation} 
which is bounded to ensure that its normal component ${\tilde{\bf
    u}_\alpha}\cdot {\bf n}$ lies between ${\tilde{\bf u}_{\alpha
    i}}\cdot {\bf n}$ and ${\tilde{\bf u}_{\alpha j}}\cdot {\bf
  n}$. The result of this is also bounded between the velocities
${{\bf u}_\alpha}_i\cdot {\bf n}$ and ${{\bf u}_\alpha}_j\cdot {\bf
  n}$.

In order to calculate the velocity between elements when using
P$_{n+1}$DG-P$_{n}$DG elements one first needs to consider that it is
also necessary to obtain the discretised non-symmetric pressure
equation which has elliptic-like properties. These properties mean it
is necessary to propagate information in all directions across the
domain, therefore taking the upwind velocity, as commonly done within
an element, is not the best strategy.

For two neighbouring CV's $i$ and $j$ with common FE interface,
$\tilde{\bf u}_\alpha$ can be calculated by using a volume-weighted
harmonic mean. However, the volume $\mathcal{V}$ of the control volume
and $\underline{\underline{\sigma}}_{\alpha}$ act on the velocity in a
very similar way. Hence, the velocity at the interface is expressed
as:
\begin{equation} 
  {\tilde{\bf u}_\alpha} = \frc{ \mathcal{V}_{j} {\bf n}^T \cdot
    \underline{\underline{\sigma}}_{\alpha j}\cdot{\bf n} \, {\bf
      u}_{\alpha i} + \mathcal{V}_{i} {\bf n}^T
    \cdot\underline{\underline{\sigma}}_{\alpha i}\cdot{\bf n}\, {\bf
      u}_{\alpha j}} { \mathcal{V}_{i} {\bf n}^T
    \cdot\underline{\underline{\sigma}}_{\alpha i}\cdot{\bf n} +
    \mathcal{V}_{j} {\bf n}^T
    \cdot\underline{\underline{\sigma}}_{\alpha j}\cdot{\bf n} }.
\end{equation}

\subsection{Weak Enforcement of Boundary Conditions}\label{bcs-rel-perm} 
Suitable boundary conditions for velocity can be applied by the
discontinuous formulation above as well as by the upwinding in the
continuous formulation. In fact, exactly the same boundary condition
implementations, as used in the discontinuous formulation, can be
applied across the boundaries of the domain by taking information from
just outside the domain rather than from the neighbouring elements.

Saturation boundary conditions are relatively straightforward. One
typically takes the saturation from just outside the domain
($S_{\alpha, \text{bc}}$) and the flux becomes $\left({\bf n}\cdot
{\bf u}_{\alpha}\right)S_{\alpha,\text{bc}}$, if ${\bf n}\cdot {\bf
  u}_{\alpha} < 0$. If the velocity is also specified (${\bf
  u}_{\alpha,\text{bc}}$) then this becomes part of the incoming flux
$\left({\bf n}\cdot {\bf
  u}_{\alpha,\text{bc}}\right)S_{\alpha,\text{bc}}$.

In case of specified pressure for outflow boundaries
$\left(\right.{\bf n}\cdot {\bf u}_\alpha>0\left.\right)$ the flux is
${\bf n}\cdot {\bf u}_\alpha {S_\alpha}$, in which neither ${\bf
  u}_\alpha$ nor ${S_\alpha}$ are specified.

However, for inflow boundaries with specified pressure
$\left(\right.{\bf n}\cdot {\bf u}_\alpha<0\left.\right)$ the flux is
${\bf n}\cdot {\bf u}_{\alpha,\text{bc}} S_{\alpha,\text{bc}}$, in
which ${\bf u}_{\alpha,\text{bc}}$ is calculated using:
\begin{equation}
  {\bf u}_{\alpha,\text{bc}} = 
    \underline{\underline{\sigma}}_{\alpha,\text{bc}} ^{-1}
    \underline{\underline{\sigma}}_{\alpha}  
              {\bf u}_{\alpha},
\end{equation}
where $\underline{\underline{\sigma}}_{\alpha,\text{bc}}$ is
calculated using $S_{\alpha,\text{bc}}$.

It should be noted that this flux condition must be used in both the
discretised saturation and global continuity equations, since the
latter is a summation of the former. The specified pressure condition
becomes a surface flux condition in the force balance equation
(Eqn.~\ref{force-semi-disc}), which effectively relaxes the pressure
$p$ to its face value counterpart $p_\text{bc}$ at the boundaries.

\subsection{Solving the Linear Equations}\label{Section:SolvingLinearEqns}
The global mass balance equation (Eqn.~\ref{glob-cty-matrix}) and
force balance equations (Eqn.~\ref{force-balance-matrix-form}) are
solved by substituting $\underline{\mathbf v}$ and solving the system
of equations for pressure. At time level $n+1$,
Eqns.~\ref{force-balance-matrix-form} and ~\ref{glob-cty-matrix} can
be rewritten as:
\begin{displaymath}
  \begin{cases}
    \mathbf{M} \underline{\bf v}^{n+1} = \mathbf{C}\underline{\bf
      p}^{n+1} + \underline{\bf s}_{u}^{n+1} \\ \mathbf{B}^{T}
    \underline{\bf v}^{n+1} = \underline{\bf s}_{p}^{n+1},
  \end{cases}
\end{displaymath}
respectively. Application of a discontinuous FEM for
$\mathbf{v}$ leads to a block-diagonal $\mathbf{M}$ matrix that
can be readily inverted, each block being local to an
element. Multiplying Eqn.~\ref{force-balance-matrix-form} by
$\mathbf{B}^{T}\mathbf{M}^{-1}$ and summing up with
Eqn.~\ref{glob-cty-matrix}, $\underline{\mathbf v}^{n+1}$
vanishes and the pressure equation is obtained:
\begin{equation}
  \mathbf{B}^{T}\mathbf{M}^{-1} \mathbf{C} \underline{\bf p}^{n+1} =
  \underline{\bf s}_{p}^{n+1} - \mathbf{B}^{T} \mathbf{M}^{-1}
  \underline{\bf s}_{u}^{n+1}.
  \label{final_projection}
\end{equation}
Eqn.~\ref{final_projection} is solved for pressure and then the
velocity is obtained via Eqn.~\ref{force-balance-matrix-form}. The
computationally demanding effort to solve the pressure matrix equation
arising from the fully discontinuous FEM formulation is achieved using
a multigrid-like approach \citep[see][]{Brandt}. In this case the
`fine' mesh is the one obtained by the discontinuous system, whereas
the `coarse' mesh is obtained by creating a continuous system (from
the discontinuous system) by collapsing the pressure nodes. As in
multigrid, the coarse mesh solution is used to accelerate the
convergence of the fine mesh solution by calculating the error of the
current approximation.

\section{Model set-up}\label{setup}
The numerical method is evaluated using four test cases (\ref{classical_BL} --
\ref{res4}). All simulate two-phase flow. The Brooks-Corey model
\citep{brooks_1964}: 
\begin{eqnarray}
k_{rw}\left ( S_{w} \right ) &=& \left ( \frac{S_w-S_{wirr}}{1-S_{wirr}-S_{nwr}} \right )^{n_w}, \\
k_{rw}\left ( S_{w} \right ) &=& \left ( \frac{S_{nw}-S_{nwr}}{1-S_{wirr}-S_{nwr}} \right)^{n_{nw}}, 
\end{eqnarray}
where $S_w$ and $S_{nw}$ are the wetting and non-wetting phase
saturations, respectively, is used to calculate the relative
permeabilities for all test cases. The length of all computational
domains is 1 non-dimensional length unit. Computational domain height,
boundary conditions, exponents for the wetting $n_w$ and non-wetting
$n_{nw}$ phases in the Brooks-Corey model, along with the irreducible
wetting phase saturation $S_{wirr}$, residual non-wetting phase
saturation $S_{nwr}$ and mobility $M^0$ for each test case are shown
in Table~\ref{tab:1}.

\begin{table}[h!]
  \centering
  \small{
    \caption{Model set-up for test cases \ref{classical_BL} --
      \ref{res4}.\label{tab:1}}
    \begin{tabular}{c c c c c c c c c c c c c c c c } 
      \hline
      & $M^0$ & $\phi$ & $\mathbf{K}_\text{1}$ & $\mathbf{K}_\text{2}$ & $S_{wirr}$ & $S_{nwr}$ & $n_{w}$ & $n_{nw}$ & $u_{\mathrm{in}}$ & $\Delta p$ & height \\ \hline
      M1 & 10 & 0.2 & 1.0 & N/A    & 0.2 & 0.3 & 2 & 2 & 0.2 & N/A & N/A\\ 
      M2 & 10 & 0.4 & 1.0 & 2.5    & 0.1 & 0.2 & 2 & 2 & 1.0 & N/A & 1.0\\ 
      M3 & 10 & 0.5 & 0.5 & 10$^4$ & 0.1 & 0.3 & 2 & 2 & 1.0 & N/A & 1.0\\ 
      M4 & 10 & 0.5 & 1.0 & 10$^2$ & 0.0 & 0.0 & 2 & 2 & N/A & 1.0 & 0.1\\ \hline
    \end{tabular}
  }
\end{table}

In all test cases, the domain is initially saturated with the
non-wetting phase at $(1- S_{wirr}).$ The wetting phase is then
injected through the left boundary, displacing the non-wetting phase
towards the opposite boundary. The wetting phase is injected at a
uniform, constant velocity ($u_{\mathrm{in}}$) from the left boundary
for test cases \ref{res2} -- \ref{res3}. In test case \ref{res2} the
inlet velocity is weighted by the permeability. For the last test case
(\ref{res4}), the flow is driven by a pressure difference ($\Delta p$)
between the inflow (left) and outflow (right) boundaries of the
computational domain. The pressure level at the outflow boundary is
set to 0 for all test cases.

For the test cases considered here, the porosity $\phi$ is assumed
uniform and for case \ref{classical_BL}, $\mathbf{K}$ is also
uniform. However, for cases \ref{res2} -- \ref{res4}, $\mathbf{K}$ is
non-uniform and regions of constant $\mathbf{K}$ are used. For all
cases, $\mathbf{K}$ is isotropic. The subscript in $\mathbf{K}$
denotes region identities (see Figs.~\ref{fig:4reg_BL_schematic},
\ref{fig:square_permeability} and \ref{fig:wedge_permeability}).

\section{Results}\label{res}
\subsection{The Buckley-Leverett test case}\label{classical_BL}

The Buckley-Leverett problem \citep[non-wetting phase displaced by a
  wetting phase,][]{buckley1942} is simulated here.  Despite the
one-dimensional (1D) nature of the problem, 2D and three-dimensional
(3D) simulations are performed to evaluate the multi-dimensional
capabilities of the model (see Fig.~\ref{fig:maps2d_3d}). The
\PN[1]{1}, \PN[1]{2} and \PN[2]{1}DG element pairs are used for these
simulations.

\begin{figure}[h!]
  \begin{center}
    \includegraphics[width=0.75\textwidth]{BL_3D_merged}
    \caption{Wetting phase volume fraction for the Buckley-Leverett
      test case using the \PN[1]{2} element pair in 2D (a) and 3D
      (b). \label{fig:maps2d_3d}}
  \end{center}
\end{figure}

\begin{figure}[h!]
  \begin{center}
    \vbox{\hbox{
        \hspace{0.25cm}\includegraphics[width=0.45\textwidth]{P1DGP1_M10.eps} 
        \hspace{0.25cm}\includegraphics[width=0.45\textwidth]{P1DGP2_M10.eps}}
      \hbox{
        \vspace{-0.cm}\hbox{\hspace{2.5cm}(a)} 
        \vspace{-0.cm}\hbox{\hspace{6.5cm}(b)}}
      \vspace{0.0cm}\hbox{
        \hspace{0.5cm}\includegraphics[width=0.45\textwidth]{P2DGP1DG_M10.eps}
        \hspace{0.5cm}\includegraphics[width=0.45\textwidth]{3D_three_elements.eps}}
      \hbox{
        \vspace{-0.cm}\hbox{\hspace{2.5cm}(c)} 
        \vspace{-0.cm}\hbox{\hspace{6.5cm}(d)}}}
    \caption{Wetting phase volume fraction along at the geometry
      centre-line at non-dimensional time $t=0.2$ for the
      Buckley-Leverett test case.  2D results with the \PN[1]{1},
      \PN[1]{2} and \PN[2]{1}DG element pairs are shown in (a), (b)
      and (c), respectively.  3D results for an unstructured mesh are
      shown in (d). The 1D analytic solution is shown for comparison
      in all plots.\label{fig:BL_tests}}
  \end{center}
\end{figure}

\begin{figure}[h!]
  \begin{center}
    \includegraphics[width=0.70\textwidth]{BL_2D_Errors.eps}
    \caption{Error metrics versus number of elements for the
      Buckley-Leverett test case for element pairs \PN[1]{1},
      \PN[1]{2} and \PN[2]{1}DG in 2D. \label{fig:errors_BL_2D}}
  \end{center}
\end{figure}

Fig.~\ref{fig:BL_tests} (a), (b) and (c) shows phase volume fraction
profiles of the wetting phase for three meshes in 2D at
non-dimensional time $t=0.2$. All meshes are structured and have 4
elements perpendicular to the flow and 120, 240 or 480 elements
parallel to the flow (see Fig.~\ref{fig:maps2d_3d}, a). The geometry
centre-line is used to plot these profiles. All element pairs
satisfactorily capture the discontinuity in the phase volume fraction
and as the number of elements is increased the error is reduced. Due
to this discontinuity in the solution variables, a first-order
convergence rate is achieved even when the high-order element pair is
used (Fig.~\ref{fig:errors_BL_2D}). This is because high-order methods
switch to simple upwinding at the discontinuity.

Fig.~\ref{fig:BL_tests} (d) shows results obtained for a
Buckley-Leverett test case in 3D. A 1207-element fully unstructured
mesh is used for these simulations. The geometry centre-line is used
to plot these profiles. All the element pairs are again capable of
accurately capturing the discontinuity in the phase volume fraction.

\subsection{Immiscible Displacement in Heterogeneous Porous Media}\label{res2}

This test case is designed to demonstrate the numerical robustness of
the method for modelling heterogeneous porous media and it is loosely
based on the physical experiment presented in \cite{dawe_2008}. As
shown in Fig.~\ref{fig:4reg_BL_schematic} the permeability field is
non-uniform.

\begin{figure}[h!]
  \begin{center}
    \includegraphics[width=0.6\textwidth]{4Rregion_BLv2}
    \caption{Schematic of the computational domain along with the
      material properties and boundary conditions for the
      heterogeneous permeability test case. Darker areas
      ($\mathbf{K}_\text{1}$) represent regions with high
      permeability. \label{fig:4reg_BL_schematic}}
  \end{center}
\end{figure}

\begin{figure}[h!]
  \vbox{\hbox{
      \hspace{-1.8cm}\includegraphics[width=0.6\textwidth]{CG_coarse083.png}
      \hspace{-0.2cm}\includegraphics[width=0.6\textwidth]{CG_fine083.png}}
    \vspace{1.cm}\hbox{
      \hspace{-1.8cm}\includegraphics[width=0.6\textwidth]{DG_coarse083.png}
      \hspace{-0.2cm}\includegraphics[width=0.6\textwidth]{DG_fine083.png}}}
  \caption{Wetting phase volume fraction maps at time $t=0.16$ for the
    heterogeneous permeability test case. Top row: \PN[1]{2}
    elements. Bottom row: \PN[2]{1}DG elements. Left column: coarse
    mesh. Right column: fine mesh. The meshes used for these
    simulations are also shown here.\label{fig:4reg_maps}}
\end{figure}

Two sets of simulations are performed using coarse (402 elements) and
fine (3714 elements) meshes; in the first set the \PN[1]{2} element
pair is used, while in the second one the \PN[2]{1}DG element pair is
used. The time step sizes for the coarse and fine mesh simulations are
5$\times$10$^{-3}$ and 1$\times$10$^{-3}$ non-dimensional time units,
respectively.

Fig.~\ref{fig:4reg_maps} shows the wetting phase volume fraction
maps (at time $t=0.16$) for all four simulations. The results yielded
by the experiment (a) are depicted in the top left, coarse mesh, and
top right, fine mesh. The results obtained from experiment (b) located
on the bottom left, coarse mesh, and bottom right, fine mesh.

Fig.~\ref{fig:4reg_maps} shows the wetting phase volume fraction maps
(at time $t=0.16$) for all four simulations. Results from the
\PN[1]{2} element pair set of simulations is shown in the top row.
Results from the \PN[2]{1}DG element pair set of simulations is shown
in the bottom row.

A physical experiment with similar set-up was performed by
\citet{dawe_2008} to investigate miscible and immiscible displacement
in heterogeneous permeability and wettability cases. For immiscible
displacement, the wetting phase was uniformly injected at rate of
$\approx 1$ml/min in a domain of 20cm$\times$10cm$\times$0.6cm. Glass
ballotini beads produced porosity of 0.4. The permeability ratio was
2.5.

Snapshots in Fig.~\ref{fig:4reg_maps} show the wetting phase flood
`fingering' across the four region intersection, demonstrating the
preferential flow through highly permeable matrix. This is in good
qualitative agreement with experimental results from \cite{dawe_2008}
\citep[see Fig.~\ref{fig:dawe_real} taken from ][]{dawe_2008}.

\begin{figure}[h!]
  \begin{center}
    \includegraphics[width=0.75\textwidth]{Real_quadrant}
    \caption{Wetting phase volume fraction maps from the experiments
      of \cite{dawe_2008}. \label{fig:dawe_real}}
  \end{center}
\end{figure}

\subsection{Underground Caisson}\label{res3}

This test case is similar to that in Section~\ref{res2}, however, the
permeability ratio is considerably higher at $10^4$. The computational
domain considered here is characterised by a low-permeability square
region of length 0.5 non-dimensional units, which is embedded in a
high-permeability square domain.

\begin{figure}[h!]
  \vbox{
    \hbox{
      \hspace{3.5cm} \includegraphics[width=0.45\textwidth]{square_permeability_lowv2}}
    \vspace{0.25cm}
    \hbox{
      \hspace{0.4cm}\includegraphics[width=0.45\textwidth]{cg_square_015_low}
      \hspace{0.4cm}\includegraphics[width=0.45\textwidth]{cg_square_05_low}}
    \vspace{0.25cm}
    \hbox{
      \hspace{0.4cm}\includegraphics[width=0.45\textwidth]{dg_square_015_low}
      \hspace{0.4cm}\includegraphics[width=0.45\textwidth]{dg_square_05_low}}}
  \caption{Schematic of the computational domain along with the
    permeability map and mesh used for the underground caisson test
    case (top).  Wetting phase volume fraction maps at non-dimensional
    times $t=0.15$ (left) and $t=0.5$ (right). \PN[1]{2} element pair
    results are shown in the middle row. \PN[2]{1}DG element pair
    results are shown in the bottom row.
    \label{fig:square_permeability}}
\end{figure}

Two numerical simulations are performed. The first one uses the
\PN[1]{2} element pair and a time step size of 5$\times$10$^{-3}$
non-dimensional time units and the second one uses the \PN[2]{1}DG
element pair and a time step size of 1$\times$10$^{-4}$. The mesh and
permeability map for these simulations is shown in
Fig.~\ref{fig:square_permeability} (top).

Fig.~\ref{fig:square_permeability} shows snapshots of the wetting
phase volume fraction field at non-dimensional times $t = 0.15$ and
0.5 for both simulations.  Numerical dispersion can be seen in the
solution of the \PN[1]{2} element pair simulation. This is because
CV's span material property interfaces. However, this is not the case
for the \PN[2]{1}DG element pair and the wetting phase here does not
diffuse into the low-permeability region of the computational domain.
Despite this small difference, both simulations are in good agreement
and provide satisfactory results with respect to dispersing the
wetting phase into the low-permeability region.

\subsection{High-Permeability high-aspect ratio region}\label{res4}

Fractures and cracks in porous media are ubiquitous, however, they are
not easy to model due to their high aspect ratio and small size
compared to the reservoir domain. Nonetheless, cracks may have a
dramatic impact in the overall behaviour of the flow as they can
channelise the flow through them. Usually, they are characterised by a
thin tip and a high permeability. A simplified 2D wedge-shaped
high-permeability region embedded in a rectangular low-permeability
domain is considered here. The permeability map is shown in
Fig.~\ref{fig:wedge_permeability}. The wedge aspect ratio is $1/30$
and its maximum height is 0.025.

\begin{figure}[h!]
  \vbox{
    \hbox{
      \hspace{-0.cm}\includegraphics[width=1.0\textwidth]{wedge_permeabilityv2}}
    \vspace{-1.cm}\hbox{\hspace{4cm}(a)}
    \vspace{.5cm}
    \hbox{
      \hspace{-0.cm}\includegraphics[width=1.0\textwidth]{wedge_0011}}
    \vspace{-0.cm}\hbox{\hspace{4cm}(b)}
    \vspace{.5cm}
    \hbox{
      \hspace{-.cm}\includegraphics[width=1.0\textwidth]{wedge_0014}}
    \vspace{-0.cm}\hbox{\hspace{4cm}(c)}
  }
  \caption{Schematic of the computational domain along with the
    permeability map for the high-permeability wedge-shaped region
    test case (a). Snapshots of the wetting phase volume fraction
    field for a fine mesh \PN[1]{2} simulation (top), coarse mesh
    \PN[1]{2} simulation (middle) and coarse mesh \PN[2]{1}DG
    simulation (bottom) at non-dimensional times $t = 0.011$ (b) and
    $t=0.0014$ (c). The meshes used for these simulations are also
    shown in (b) and (c).
    \label{fig:wedge_permeability}}
\end{figure}

Three numerical simulations are performed in total. Two use the
\PN[1]{2} element pair and one uses the \PN[2]{1}DG element pair.
\PN[1]{2} element pair simulations are performed using a coarse (672
elements) and a fine (56902 elements) mesh. The coarse mesh is used
for the \PN[2]{1}DG element pair simulation.

Fig.~\ref{fig:wedge_permeability} shows wetting phase volume fraction
maps at non-dimensional times $t=0.011$ and 0.014. The \PN[1]{2}-based
solutions disperse the wetting phase into the low-permeability
matrix. This is more pronounced for the coarse mesh simulation. It can
be seen how the front of the wetting phase is considerably delayed
compared with the other two simulations. On the other hand, the
\PN[2]{1}DG solution does not disperse the wetting phase into the
matrix and even a very coarse mesh is sufficient to accurately capture
the flow.

\section{Concluding Remarks}\label{conc}
A novel, high-order, control-volume-finite-element (CVFE) method with
$n^{\rm th}$-order representation for velocity and $m^{\rm th}$-order
for pressure is presented. The method conserves mass and ensures that
the extended Darcy equations for multiphase flow are exactly
enforced. Moreover, for \PN[n+1]{n}DG element pairs control volumes
do not span elements; this resolves a long-standing problem associated
with the use of classical CVFE methods in highly heterogeneous porous
media flow modelling. In addition, the use of force per unit volume
instead of Darcy velocity significantly improves the computational
cost and memory usage.

The new formulation is evaluated using the classical Buckley-Leverett
benchmark and first-order convergence rates are achieved, which is not
common in the literature \citep[see][]{Schmid,Hoteit}. More numerical
simulations are performed to demonstrate the capabilities of the
methods presented under heterogeneous and high aspect ratio
computational domains, showing that this method is able to resolve
successfully the flow under those circumstances. Furthermore, the
\PN[n+1]{n}DG element pair shows that it completely removes numerical
dispersion induced by heterogeneous permeabilities.

The formulation presented here is implemented in the open-source
software framework
\href{http://www3.imperial.ac.uk/earthscienceandengineering/research/amcg/fluidity}{Fluidity}
for multiphase flows and it has been extended to include/model
multi-component, interface-capturing
\citep{pavlidis_2013b,pavlidis_2014,xie_2014}, segregation in granular
flows~\citep{percival_2014} and oil reservoirs
\citep{jackson_2013,mostaghimi_2015, Salinas}.

\ack Dr D. Pavlidis would like to acknowledge the support from the
following research grants: EPSRC ($\lq$Computational Modelling for
Advanced Nuclear Power Plants') and EU/FP7 THINS. Funding for Dr
P. Salinas from ExxonMobil is gratefully acknowledged. Dr Z. Xie is
supported by EPSRC ($\lq$Multi-Scale Exploration of Multiphase Physics
in Flows' -- MEMPHIS). Part funding for Prof Jackson under the TOTAL
Chairs programme at Imperial College is also acknowledged.

%% The Appendices part is started with the command \appendix;
%% appendix sections are then done as normal sections
%% \appendix

%% \section{}
%% \label{}

%% References with bibTeX database:
\bibliographystyle{elsarticle-harv}
\bibliography{references}

\end{document}

%%
%% End of tex file.