### 2018-04-15 Papers/AppliedMaths2018/Highlights.tex
Initial submission of final manuscript 15th April 2018

\documentclass[a4paper,11pt]{article}

\newcommand{\ie}{{\it i.e., }}
% define the title
\author{B.~Lashore}
\title{Highlights}
\begin{document}
% generates the title
\maketitle
% insert the table of contents
%\tableofcontents
%\section{Some Interesting Words}

\begin{enumerate}
  \item Performed a model order reduction (\it upscale) of a permeability field using singular value decomposition (SVD)
  \item Reduced order model (ROM) for SVD was implimented in the principal component space (PCS)
  \item Benchmarked the SVD model by comparing with Arithmetic mean, Harmonic mean and probability density function (PDF) models
  \item A novel high-order accurate control volume FEM multiphase flow simulator also compared the simulations from all the models
  \item The SVD model presented a better representation compared to the other models 
\end{enumerate}
\end{document}