### 2018-12-06 Papers/AppliedMaths2018/Resubmission/Highlights_resubmission.tex
Corrected typo and re-committed

\documentclass[a4paper,11pt]{article}

\newcommand{\ie}{{\it i.e., }}
% define the title
\author{B.~Lashore}
\title{Highlights}
\begin{document}
% generates the title
\maketitle
% insert the table of contents
%\tableofcontents
%\section{Some Interesting Words}

\begin{enumerate}
  \item Coupled linear regression with Principal Component Analysis (PCA) using SVD, to obtain a 2D reduced order model
  \item Benchmarked the 2D PCA and linear regression method by comparing with Arithmetric mean, Harmonic mean and PDF models
  \item Demonstrated that the PCA-Linear regression technique can be extended to 3D models with Higher Order SVD (HOSVD)
  \item Performed a qualitative comparision of the models and simulation results for both the 2D and the 3D cases
  \item The coupled PCA and linear regression model model presented a better representation compared to the other models
\end{enumerate}
\end{document}